%!TEX root = ./ft-hd.tex 
\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{times}
%\usepackage[small, compact]{titlesec}
\usepackage{subfigure}
\usepackage{color}
\usepackage{url}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{mathtools}
\textheight 9.05in 
\textwidth 6.7in 

\usepackage{tikz}
\usetikzlibrary{positioning, calc, chains}

% Complex \xxx for making notes of things to do.  Use \xxx{...} for general
% notes, and \xxx[who]{...} if you want to blame someone in particular.
% Puts text in brackets and in bold font, and normally adds a marginpar
% with the text ``xxx'' so that it is easy to find.  On the other hand, if
% the comment is in a minipage, figure, or caption, the xxx goes in the text,
% because marginpars are not possible in these situations.
{\makeatletter
 \gdef\xxxmark{%
   \expandafter\ifx\csname @mpargs\endcsname\relax % in minipage?
     \expandafter\ifx\csname @captype\endcsname\relax % in figure/caption?
       \marginpar{xxx}% not in a caption or minipage, can use marginpar
     \else
       xxx % notice trailing space
     \fi
   \else
     xxx % notice trailing space
   \fi}
 \gdef\xxx{\@ifnextchar[\xxx@lab\xxx@nolab}
 \long\gdef\xxx@lab[#1]#2{{\bf [\xxxmark #2 ---{\sc #1}]}}
 \long\gdef\xxx@nolab#1{{\bf [\xxxmark #1]}}
 % This turns them off:
%\long\gdef\xxx@lab[#1]#2{}\long\gdef\xxx@nolab#1{}%
}


\newcommand{\head}[2]{#1_{[#2]}}
\newcommand{\tail}[2]{#1_{\overline{[#2]}}}
\newcommand{\abs}[1]{|#1|}
\newcommand{\ceil}[1]{\lceil#1\rceil}
\newcommand{\floor}[1]{\lfloor#1\rfloor}
\newcommand{\inner}[1]{\langle#1\rangle}
\newcommand{\tnorm}[2]{\left \lVert#2\right \rVert_{#1}}
\newcommand{\norm}[2]{\lVert#2\rVert_{#1}}
\newcommand{\range}[3]{#1_{\overline{[#2]}\cap[#3]}}
%\newcommand{\I}{\mathrm{I}}
\newcommand{\gaplinf}{\mathsf{Gap}\ell_{\infty}}
\newcommand{\gaplinfL}{\mathsf{LooseGap}\ell_{\infty}}
\newcommand{\multiLinf}{\mathsf{Multi}\ell_{\infty}}
\newcommand{\indlinf}{\mathsf{Ind}\ell_{\infty}}
\newcommand{\indlinfL}{\mathsf{LooseInd}\ell_{\infty}}
\newcommand{\IC}{\mathrm{IC}}
\newcommand{\wh}{\widehat}
\newcommand{\twidt}[1]{\tilde{#1}^{(t)}}
\newcommand{\twid}[1]{\tilde{#1}}
\newcommand{\ostar}{O^*}
\newcommand{\cc}{\mathcal{C}}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator*{\median}{median}
\DeclareMathOperator*{\mean}{mean}
%\DeclareMathOperator*{\E}{\mathbb{E}}
%\DeclareMathOperator*{\Q}{Quant}
\newcommand{\Q}[1]{\operatorname*{quant^{\mathnormal{#1}}}}
\def\R{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\C{\mathbb{C}}
\def\F{\mathcal{F}}
\def\B{\mathbb{B}}
\def\eps{\epsilon}



\newcommand{\etal}{{\em et al.}}
%\newcommand{\qed}{\mbox{}\hspace*{\fill}\nolinebreak\mbox{$\rule{0.6em}{0.6em}$}}
\newcommand{\expect}{{\bf \mbox{\bf E}}}
\newcommand{\prob}{{\bf \mbox{\bf Pr}}}
\newcommand{\opt}{{\rm OPT}}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\newcommand{\e}{{\epsilon}}
\newcommand{\A}{{\mathcal{A}}}
\newcommand{\E}{{\mathcal{E}}}

\newcommand{\comment}[1]{{\color{gray}[\textsf{#1}]}}
\newcommand{\grainofsalt}[1]{{\color{gray}[\textsf{#1}]}}
\newcommand{\redospace}{\small\renewcommand{\baselinestretch}{1.5}\normalsize}
\newcommand{\undospace}{\small\renewcommand{\baselinestretch}{1}\normalsize}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{proposition}[theorem]{Proposition}
%\newenvironment{proof}{{\bf Proof:}}{$\qed$\par}
%\newenvironment{proofnoqed}{{\bf Proof:}}{}
\newenvironment{proofof}[1]{\noindent{\bf Proof of #1:}}{$\qed$\par}
%\newenvironment{proofsketch}{{\sc{Proof Outline:}}}{$\qed$\par}
\usepackage{hyperref}
\hypersetup{
bookmarksnumbered
}


\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\newcommand{\I}{\mathcal{I}}

\renewcommand{\b}{{\mbox {\bf b}}}
\newcommand{\barb}{{\mbox {\bf \bar{b}}}}
\newcommand{\coll}{{\mathcal R}}
\newcommand{\colltwo}{{\mathcal T}}
\newcommand{\collpart}{{\mathcal S}}
\newcommand{\edgesets}{\mathcal X}

\newcommand{\sinc}{\text{sinc}}
\DeclareMathOperator{\err}{Err}
\newcommand{\poly}{\text{poly}}
\newcommand{\SNR}{\text{SNR~}}
%\renewcommand{\E}{{\mathcal{E}}}
\newcommand{\SCALE}{\text{SCALE}}
\renewcommand{\top}{\text{top}}

\newcommand{\Ev}{\mathcal{E}}
\newcommand{\one}{\mathbf{1}}
\newcommand{\csl}{(n/B)^{1/d}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\fc}{F}
%\newcommand{\gl}{GL(d, \Z_n)}
\newcommand{\gl}{\mathcal{M}_{d\times d}}
\newcommand{\rect}{\text{rect}}
\newcommand{\spc}{12}%sparsity of z is |S|/\log^\spc n
\newcommand{\gapc}{4}%gap

\renewcommand{\H}{\mathcal{W}}
\newcommand{\h}{\mathbf{w}}

\newcommand{\zero}{\mathbf{0}}

\newcommand{\defsk}{Let $S\subseteq \nsq, |S|\leq 2k$, be such that $||x_{\nsq\setminus S}||_\infty\leq \mu$. Suppose that $||x||_\infty/\mu\leq N^{O(1)}$.~Let $B\geq \GT\cdot k/\alpha^d$.~}
%\newcommand{\defskmin}{Let $S\subseteq \nsq, |S|\leq 2k$. ~Let $B\geq \GT\cdot k/\alpha^d$.~}


\newcommand{\f}{\mathbf{f}}
\newcommand{\quant}{\text{quant}}
\newcommand{\GS}{(2\pi)^{d\cdot \fc}}
\newcommand{\GSS}{(2\pi)^{2d\cdot \fc}}
\newcommand{\GT}{(2\pi)^{4d\cdot \fc}}
\newcommand{\GI}{(2\pi)^{-d\cdot \fc}}
\begin{document}

\begin{titlepage}
\title{Sparse Fourier Transform in Any Constant Dimension with Nearly-Optimal Sample Complexity in Sublinear Time}
\author{Michael Kapralov\\EPFL}

\maketitle

\begin{abstract}
We consider the problem of computing a $k$-sparse approximation to the Fourier transform of a length $N$ signal. Our main result is a randomized algorithm for computing such an approximation (i.e. achieving the $\ell_2/\ell_2$ sparse recovery guarantees using Fourier measurements) using $O_d(k\log N\log\log N)$ samples of the signal in time domain that runs in time $O_d(k\log^{d+3} N)$, where $d\geq 1$ is the dimensionality of the Fourier transform. The sample complexity matches the lower bound of $\Omega(k\log (N/k))$ for non-adaptive algorithms due to~\cite{DIPW} for any $k\leq N^{1-\delta}$ for a constant $\delta>0$ up to an $O(\log\log N)$ factor. Prior to our work a result with comparable sample complexity $k\log N \log^{O(1)}\log N$ and sublinear runtime was known for the Fourier transform on the line~\cite{IKP}, but for any dimension $d\geq 2$ previously known techniques either suffered from a $\poly(\log N)$ factor loss in sample complexity or required $\Omega(N)$ runtime.
\end{abstract}
\thispagestyle{empty}
\end{titlepage}

\tableofcontents

\newpage


\newcommand{\nsq}{{[n]^d}}


\input{intro.tex}
\input{preliminaries.tex}
\input{algo.tex}


\begin{algorithm}

\caption{\textsc{ReduceL1Norm}$\left(\wh{x}, \chi, k, \chi^{(t)}, k, \{m(\wh{x}, H_r, a\star (\one, \h))\}_{r=1, a\in \A_r, \h\in \H}^{r_{max}}, \nu, \mu\right)$}\label{alg:l1-norm-reduction} 
\begin{algorithmic}[1] 
\Procedure{ReduceL1Norm}{\mbox{$\wh{x}, \chi, k, \chi^{(t)}, k, \{m(\wh{x}, H_r, a\star (\one, \h))\}_{r=1, a\in \A_r, \h\in \H}^{r_{max}}, \nu, \mu$}}
\State $\chi^{(0)} \gets 0$ \Comment{in $\C^n$}
\State $B\gets  \GT\cdot k/\alpha^d$
\For{$t=0$ to $\log_2 (\log^4 N)$} 
\For {$r=1$ to $r_{max}$}
\State $L_r \gets \textsc{LocateSignal}\left(\chi+\chi^{(t)}, k, \{m(\wh{x}, H_r, a\star (\one, \h))\}_{r=1, a\in \A_r, \h\in \H}^{r_{max}}\right)$ 
\EndFor
\State $L\gets \bigcup_{r=1}^{r_{max}} L_r$
%\State Choose $\Sigma_q^{est}, b_q^{est}, a_q^{est}$ uniformly at random, $\Sigma_q^{est}$ invertible, for $q=1,\ldots, q_{max}=\Theta(\log\log n)$ 
\State $\chi' \gets \Call{EstimateValues}{\wh{x},  \chi+\chi^{(t)}, L, 4k, 1, \frac1{1000} \nu 2^{-t}+4\mu, C(\log\log N+d^2+\log (B/k))}$
\State $\chi^{(t+1)} \gets \chi^{(t)} + \chi'$
\EndFor
\State \textbf{return} $\chi+\chi^{(T)}$
\EndProcedure 
\end{algorithmic}
\end{algorithm}









\input{l1-part1.tex}
\input{l1.tex}
\input{combo.tex}
\input{linf.tex}
\input{constsnr.tex}
\input{utilities.tex}
\input{semiequi.tex}

\section{Acknowledgements}
The author would like to thank Piotr Indyk for many useful discussions at various stages of this work.


%\bibliographystyle{alpha}
%\bibliography{paper}

\newcommand{\etalchar}[1]{$^{#1}$}
\begin{thebibliography}{HIKP12b}

\bibitem[AGS03]{AGS}
A.~Akavia, S.~Goldwasser, and S.~Safra.
\newblock Proving hard-core predicates using list decoding.
\newblock {\em FOCS}, 44:146--159, 2003.

\bibitem[Aka10]{Ak}
A.~Akavia.
\newblock Deterministic sparse {F}ourier approximation via fooling arithmetic
  progressions.
\newblock {\em COLT}, pages 381--393, 2010.

\bibitem[BCG{\etalchar{+}}12]{BCGLS}
P.~Boufounos, V.~Cevher, A.~C. Gilbert, Y.~Li, and M.~J. Strauss.
\newblock What's the frequency, {K}enneth?: Sublinear {F}ourier sampling off
  the grid.
\newblock {\em RANDOM/APPROX}, 2012.

\bibitem[Bou14]{Bourgain2014}
J.~Bourgain.
\newblock An improved estimate in the restricted isometry problem.
\newblock {\em GAFA}, 2014.

\bibitem[CCFC02]{CCF}
M.~Charikar, K.~Chen, and M.~Farach-Colton.
\newblock Finding frequent items in data streams.
\newblock {\em ICALP}, 2002.

\bibitem[CGV12]{CGV}
M.~Cheraghchi, V.~Guruswami, and A.~Velingker.
\newblock Restricted isometry of {F}ourier matrices and list decodability of
  random linear codes.
\newblock {\em SODA}, 2012.

\bibitem[Cip00]{citeulike:6838680}
B.~A. Cipra.
\newblock {The Best of the 20th Century: Editors Name Top 10 Algorithms}.
\newblock {\em SIAM News}, 33, 2000.

\bibitem[CP10]{CP}
E.~Candes and Y.~Plan.
\newblock A probabilistic and ripless theory of compressed sensing.
\newblock {\em IEEE Transactions on Information Theory}, 2010.

\bibitem[CT06]{CTao}
E.~Candes and T.~Tao.
\newblock Near optimal signal recovery from random projections: Universal
  encoding strategies.
\newblock {\em IEEE Trans. on Info.Theory}, 2006.

\bibitem[DIPW10]{DIPW}
K.~{Do Ba}, P.~Indyk, E.~Price, and D.~Woodruff.
\newblock {Lower Bounds for Sparse Recovery}.
\newblock {\em SODA}, 2010.

\bibitem[Don06]{Don}
D.~Donoho.
\newblock Compressed sensing.
\newblock {\em IEEE Transactions on Information Theory}, 52(4):1289--1306,
  2006.

\bibitem[DR93]{DR93}
A.~Dutt and V.~Rokhlin.
\newblock Fast fourier transforms for nonequispaced data.
\newblock {\em SIAM J. Sci. Comput.}, 14(6):1368--1393, November 1993.

\bibitem[GGI{\etalchar{+}}02]{GGIMS}
A.~Gilbert, S.~Guha, P.~Indyk, M.~Muthukrishnan, and M.~Strauss.
\newblock Near-optimal sparse {F}ourier representations via sampling.
\newblock {\em STOC}, 2002.

\bibitem[GHI{\etalchar{+}}13]{GHIKPS}
B.~Ghazi, H.~Hassanieh, P.~Indyk, D.~Katabi, E.~Price, and L.~Shi.
\newblock Sample-optimal average-case sparse {F}ourier transform in two
  dimensions.
\newblock {\em Allerton}, 2013.

\bibitem[GL89]{GL}
O.~Goldreich and L.~Levin.
\newblock A hard-corepredicate for allone-way functions.
\newblock {\em STOC}, pages 25--32, 1989.

\bibitem[GLPS10]{GLPS}
A.~C. Gilbert, Y.~Li, E.~Porat, and M.~J. Strauss.
\newblock Approximate sparse recovery: optimizing time and measurements.
\newblock In {\em STOC}, pages 475--484, 2010.

\bibitem[GMS05]{GMS}
A.~Gilbert, M.~Muthukrishnan, and M.~Strauss.
\newblock Improved time bounds for near-optimal space {F}ourier
  representations.
\newblock {\em SPIE Conference, Wavelets}, 2005.

\bibitem[HAKI12]{HAKI}
H.~Hassanieh, F.~Adib, D.~Katabi, and P.~Indyk.
\newblock Faster {GPS} via the {S}parse {F}ourier {T}ransform.
\newblock {\em MOBICOM}, 2012.

\bibitem[HIKP12a]{HIKP2}
H.~Hassanieh, P.~Indyk, D.~Katabi, and E.~Price.
\newblock Near-optimal algorithm for sparse {F}ourier transform.
\newblock {\em STOC}, 2012.

\bibitem[HIKP12b]{HIKP}
H.~Hassanieh, P.~Indyk, D.~Katabi, and E.~Price.
\newblock Simple and practical algorithm for sparse {F}ourier transform.
\newblock {\em SODA}, 2012.

\bibitem[HKPV13]{heidersparse}
S.~Heider, S.~Kunis, D.~Potts, and M.~Veit.
\newblock A sparse {P}rony {FFT}.
\newblock {\em SAMPTA}, 2013.

\bibitem[HR16]{HR16}
I.~Haviv and O.~Regev.
\newblock The restricted isometry property of subsampled fourier matrices.
\newblock {\em SODA}, 2016.

\bibitem[IK14]{IK14a}
P.~Indyk and M.~Kapralov.
\newblock Sample-optimal {F}ourier sampling in any fixed dimension.
\newblock {\em FOCS}, 2014.

\bibitem[IKP14]{IKP}
P.~Indyk, M.~Kapralov, and E.~Price.
\newblock ({N}early) sample-optimal sparse {F}ourier transform.
\newblock {\em SODA}, 2014.

\bibitem[Iwe10]{Iw}
M.~A. Iwen.
\newblock Combinatorial sublinear-time {F}ourier algorithms.
\newblock {\em Foundations of Computational Mathematics}, 10:303--338, 2010.

\bibitem[Iwe12]{Iw-arxiv}
M.A. Iwen.
\newblock Improved approximation guarantees for sublinear-time {F}ourier
  algorithms.
\newblock {\em Applied And Computational Harmonic Analysis}, 2012.

\bibitem[KM91]{KM}
E.~Kushilevitz and Y.~Mansour.
\newblock Learning decision trees using the {F}ourier spectrum.
\newblock {\em STOC}, 1991.

\bibitem[LDSP08]{MRICS}
M.~Lustig, D.L. Donoho, J.M. Santos, and J.M. Pauly.
\newblock Compressed sensing mri.
\newblock {\em Signal Processing Magazine, IEEE}, 25(2):72--82, 2008.

\bibitem[LWC12]{LWC}
D.~Lawlor, Y.~Wang, and A.~Christlieb.
\newblock Adaptive sub-linear time fourier algorithms.
\newblock {\em arXiv:1207.6368}, 2012.

\bibitem[Man92]{Man}
Y.~Mansour.
\newblock Randomized interpolation and approximation of sparse polynomials.
\newblock {\em ICALP}, 1992.

\bibitem[PR13]{pawar2013computing}
S.~Pawar and K.~Ramchandran.
\newblock Computing a $k$-sparse $n$-length {D}iscrete {F}ourier {T}ransform
  using at most $4k$ samples and {$O(k log k)$} complexity.
\newblock {\em ISIT}, 2013.

\bibitem[PS15]{PZ15}
E.~Price and Z.~Song.
\newblock A robust sparse {F}ourier transform in the continuous setting.
\newblock {\em FOCS}, 2015.

\bibitem[RV08]{RV}
M.~Rudelson and R.~Vershynin.
\newblock On sparse reconstruction from {F}ourier and {G}aussian measurements.
\newblock {\em CPAM}, 61(8):1025--1171, 2008.

\bibitem[Sid11]{CT_CS}
Emil Sidky.
\newblock {What does compressive sensing mean for X-ray CT and comparisons with
  its MRI application}.
\newblock In {\em Conference on Mathematics of Medical Imaging}, 2011.

\end{thebibliography}


\begin{appendix}
\input{app.tex}
\end{appendix}

\end{document} 
